Think deep for reef resilience

### 2025-08-10 paper/paper_draft.qmd
Draft restructure of assessment methods section

**\[TI: Ben, best to use the same colours consistently for each management region. Blue,
yellow/orange, green and pink were used in for the map.\]**
Reef depth characteristics were extracted using 10m resolution bathymetry data from the
Great Barrier Reef Marine Park Authority (GBRMPA [-@GBRMPA2020]). For each reef, bathymetry
data (in raster format) were spatially matched with reef polygon features through geometric
intersection analysis. Depth values within each reef polygon were summarized using zonal
statistics to calculate minimum, mean, median, maximum, and standard deviation of depth
values. All depth values were converted to positive values (depths below sea level). The
median depth was used to provide an indicative depth value for each reef.

As bathymetry data are derived from satellite imagery they can be limited in their alignment
with the established GBRMPA reef outlines used in the investigation to characterise reef
locations. Quality control flags were assigned to identify potential data limitations: reefs
with no overlapping bathymetry data, reefs with minimum depth values above sea level
(indicating potential data artifacts), and reefs where bathymetry pixels covered less than
5% of the reef polygon area. These reefs with insufficient data coverage were removed from
analyses (842, or 22%, reefs of original 3806 reefs used in Bozec et al. [@bozec2025]).

**\[BG: I have updated now, does it make sense referring to the original 3806 reefs (I have
referred to this 3806 set of reefs in the first paragraph of the methods)?\]**

While data adopted for the given assessment are said to be accurate up to 15m from Mean Sea
Level [@GBRMPA2020], there are many sources of uncertainty in the depth estimates used for
this study. There is potential misalignment between GBRMPA reef outlines, as well as the
usual challenges of deriving data from satellite imagery. In the case of bathymetry,
estimates are subject to different bottom types (rock, sand, seagrass, etc) and how these
affect reflectance, the water penetration depth of satellite sensors, influence of local
conditions (e.g., turbidity and wave activity), and many other factors. The resulting depth
estimates are therefore highly uncertain themselves and are only adopted as a useful
indication rather than an accurate and precise report of reef depths across the GBR.

**\[TI: TODO cite Canonical Reefs repository\]**

## ADRIA-CoralBlox modelling

ADRIA.jl is a decision support platform that combines scenario analysis with biotic and
abiotic processes such as bleaching responses, thermal tolerance and flows of larvae between
reefs [@iwanaga2023]. CoralBlox.jl is a coral population growth and mortality model that
simulates groups of corals growing in diameter and competing for limited space over time
[@ribeirodealmeida2024]. We combine these tools to model the populations of corals across
the GBR. Details on how coral demographic processes are implemented in ADRIA - CoralBlox can
be found in the supplementary material.

### TODO: Section on scenario modelling

>>>>

**\[TI: Ben, best to use the same colours consistently for each management region. Blue,
yellow/orange, green and pink were used in for the map.\]**

## Assessment framework

An exploratory scenario modelling approach is adopted for this investigation. The method
involves exploring a wide range of plausible futures under different assumptions of
environmental (climatic) conditions and ecological factors. Exploratory scenario modelling
embraces uncertainty as an inherent characteristic of complex environmental systems
[@lempert2010] and is a core methodology underpinning the development of robust and adaptive
management strategies [@haasnoot2013].

Exploratory modelling approaches have gained favour in recent years as precise long-term
predictions of future conditions are often impossible in complex systems like coral reefs.
Multiple interacting and compounding sources of uncertainty, and a lack of broad agreement
as to their likelihood, create inherently uncertain outcomes. Rather than attempting to
forecast specific outcomes, exploratory modelling aims to systematically explore many
possible futures to identify patterns and, in management contexts, robust strategies that
perform well across those diverse conditions, as well as conditions in which such strategies
may fail. Such an exploratory approach lends itself well to the purpose of the
investigation. Specifically, to assess the role of DHW, connectivity, and depth to identify
the environmental and reef characteristics that support structural resilience.

To apply exploratory scenario modelling, the approach employed a modelling framework
involving two separate systems: ADRIA (the Adaptive Dynamic Reef Interventin Algorithms
platform) and CoralBlox. ADRIA is a multi-criteria decision support platform designed to
help reef managers address climate change challenges in coral reef systems [@iwanaga2023].
The ADRIA platform provides functionality for scenario generation and analysis as well as
providing the technical infrastructure to couple environmental and ecosystem
representations.

CoralBlox is a computationally efficient coral population model. It simulates
functional/morphological groups of corals over its life cycle, competing for available space
over time [@ribeirodealmeida2024]. The coupled framework is referred to as ADRIA-CoralBlox.
The coral demographic representation and associated model implementation are further
detailed in the supplementary material.

### Scenario modelling
